Title

### 2023-07-01 Template/main.tex
...

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

@@

% \linenumbers

@@

Such systems have been simulated because of their many applications, 
such as studying the emergence of multipartite entanglement~\cite{Andrea,Andrea_AD},
studying dissipative processes~\cite{Barreiro}
and modeling non-Markovian dynamics~\cite{Marsden}.
Among quantum systems, the simplest case is that of a qubit~\cite{chuangbook},
and withing them, the simplest class of channels that produce decoherence are
Pauli channels~\cite{geometry,Zbigniew,Davalos}. Indeed, they serve as
effective models for the noise affecting quantum devices~\cite{Flammia,Fawzi}.

To represent the algorithms implemented in quantum computers, either to
simulate a physical system or for some other purpose, one often uses a quantum
circuit~\cite{chuangbook}.  In this work, we shall also work with parametrized
quantum circuits, that is,  quantum circuits in which some of the operations
depend on variable parameters~\cite{cerezo}.  These circuits play an important
part in applications such as quantum machine learning~\cite{Benedetti} and
describing general quantum transformations dependent on parameters. Substantial
research has been devoted to enhancing the efficiency of these
circuits~\cite{Watanabe,Rasmussen}.

@@

% Intro  {{{
% \textbf{Lista de puntos:}
% \begin{itemize}
% \item \textbf{Circuit for a Pauli channel:}  Propose a circuit to simulate a Pauli channel on $n$ qubits. It works by creating a $2n$-qubit state on ancilla qubits.
% \item \textbf{Simulation:} Show the fidelities of simulating the circuit on a quantum computer for many one-qubit channels inside the tetrahedron.
% \item \textbf{For Pauli dynamical maps}: See how the circuit generalizes to parametrized channels and notice that many rotations are needed to create the $2n$-ancilla qubit state. \\
% \end{itemize}

In this section  we introduce the concept of quantum channels, focusing on a
specific type called Pauli channels.  Furthermore, we define Pauli dynamical
maps, which are curves of Pauli channels parametrized by a variable.
% }}}
\subsection{Quantum Channels} \label{subsec: Quantum Channels} % {{{

In quantum mechanics, a closed system's state is represented by a vector in a
Hilbert space $\mathcal{H}$.  The state's evolution is unitary and given by
Schrodinger's equation~\cite{Rieffel}.  However, in real-world situations,
quantum systems are usually open, which means that they interact with their
environment~\cite{Breuer}.  For instance, the system's state may become
entangled with the environment, leading to a loss of information about the
system's state over time.

@@

% }}}
\subsection{Pauli Channels}  \label{subsec: Pauli Channels} % {{{

@@

Pauli channels are some of the 
most fundamental noise models in quantum information science~\cite{Terhal}. 
Some notable examples of Pauli channels are the following:
\begin{itemize}
\item \textbf{Bit Flip Channel:} This is a channel that with probability $1-p$ leaves the qubit as it is 
and with probability $p$ it applies the $\sigma_1$ matrix 
(which flips the basis states $|0\rangle$ and $|1\rangle$ of the qubit), and
so it is given by:
\begin{align*}
\mathcal{E}(\rho) = (1-p) \rho + p \sigma_1 \rho \sigma_1.
\end{align*}
Analogous channels exist using $\sigma_3$ (called the bit flip channel,
which has a probability $p$ of adding a relative phase $\pi$ to the state) or 
using $\sigma_2$ (called the phase flip channel, 
which has a probability $p$ of 
flipping the base states and also add a relative phase $\pi$).
\item \textbf{Depolarizing channel:} 
This channel has a probability $1-p$
of doing nothing to the qubit and a probability $p$ of converting it into the maximally mixed state $\dfrac{1}{2} \mathbb{I}$
and it can be written as:
\begin{equation}
\mathcal{E}(\rho) = (1-p)\rho + p \dfrac{1}{2} \mathbb{I} = \left(1 - \dfrac{3p}{4} \right) \sigma + \dfrac{p}{4} \sigma_1 \rho \sigma_1 + \dfrac{p}{4} \sigma_2 \rho \sigma_2+ \dfrac{p}{4} \sigma_3 \rho \sigma_3.
\end{equation} 
\end{itemize}

@@

These conditions imply that $(\tau_1,\tau_2,\tau_3)$
has to be inside a tetrahedron with
vertices $(1,1,1), (1,-1,-1),(-1,1,-1)$ and $(-1,-1,1)$. 
Therefore, the $\tau_{1,2,3}$ are numbers between 
$-1$ and $1$, which means that the components
$r_{\alpha}$ of the density matrix are always 
attenuated and possibly sign flipped.

@@

As in the one qubit case, Pauli channels for $N$ qubits 
attenuate the components $r_{\vec{\alpha}}$ of the density matrix.
This can be seen by substituting \eref{eq:  Density matrix Nqbit} in \eref{eq: pauli channel N-qubit}
and using the property of \eref{eq: propiedad-pauli}:
\begin{align*}
\mathcal{E}(\rho) &= \dfrac{1}{2^N} \sum_{\vec{\gamma},\vec{\alpha}} k_{\vec{\gamma}} r_{\vec{\alpha}} \sigma_{\vec{\gamma}} \sigma_{\vec{\alpha}} \sigma_{\vec{\gamma}} \\
& = \dfrac{1}{2^N} \sum_{\vec{\alpha}} \left( \sum_{\vec{\gamma}} {(A^{\otimes^N})}_{\vec{\alpha},\vec{\gamma}} \; k_{\vec{\gamma}} \right) r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
\end{align*}
which means that applying the Pauli channel multiplies the 
components $r_{\vec{\alpha}}$ by $\tau_{\vec{\alpha}} := \sum_{\vec{\gamma}} {(A^{\otimes^N})}_{\vec{\alpha},\vec{\gamma}} k_{\vec{\gamma}}$.
% }}}
\subsection{Pauli Dynamical Maps} % {{{
\label{subsec: Pauli Dynamical Maps}

@@

In this section we propose a quantum circuit that simulates $N$-qubit Pauli
channels.  Moreover, we implement the circuit for $N=1$ on a quantum computer
and analyze the results using the diamond norm.
We find that close to the depolarizing channel, the general circuit simulator 
can implement such channels with the highest fidelity. 
% produce the highest 
% }}} 
\subsection{Description of the Circuit for a Pauli Channel} %{{{
\label{subsec: Description of the circuit}

% To design the circuit, we recall that a Pauli channel, as the one in
% \eref{eq: pauli channel N-qubit}, applies each Pauli operator
% $\sigma_{\vec{\gamma}}$ on $\rho$ with a probability $k_{\vec{\gamma}}$.
% Given this, it is easy to construct a quantum circuit that implements 
% a Pauli channel with the help of $2N$ ancilla qubits,
% see \fref{fig: circuit-pauli}.
% The general idea is to produce the probabilities $k_{\gamma}$ by creating the state
% \begin{eqnarray}
% \label{eq: state0}
% \sum_{\vec{\gamma}} b_{\vec{\gamma}} |\vec{\gamma} \rangle
% \end{eqnarray}
% in the ancilla qubits, 
% where $b_{\vec{\gamma}}$ are numbers such that $
% |b_{\vec{\gamma}}|^2 = k_{\vec{\gamma}}$.
% The $2N$-qubit state $|\vec{\gamma}\rangle$  is defined as $|\gamma_1 \rangle
% \cdots |\gamma_N\rangle$.

% {\color{orange} Cambiar por:  
To design the circuit that implements \eref{eq: pauli channel N-qubit}, we construct a state that 
includes the probabilities $k_{\gamma}$ on the ancilla qubits and subsequently
apply controlled Pauli operations on the main qubits.  The circuit that does
this is presented in \fref{fig: circuit-pauli}.

\begin{figure} % {{{
\centering
\includegraphics{images/circuito_general.pdf}
\caption{{\bf Circuit for an $N$-qubit
Pauli channel on $N$ qubits.}
The circuit creates the state $\sum_{\vec{\gamma}} b_{\vec{\gamma}}|\vec{\gamma}\rangle$ 
on the $2N$ ancilla qubits denoted by $\tilde{q}$. }
\label{fig: circuit-pauli}
\end{figure} % }}}

The first part of the circuit involves the creation of the state
\begin{eqnarray}
\label{eq: state}
\text{Ancilla state =} 
\sum_{\vec{\gamma}} b_{\vec{\gamma}} |\vec{\gamma} \rangle
\end{eqnarray}
on the ancilla qubits, where $b_{\vec{\gamma}}$ are numbers such that $
{|b_{\vec{\gamma}}|}^2 = k_{\vec{\gamma}}$ and
the $2N$-qubit state $|\vec{\gamma}\rangle$ is defined as $|\gamma_1\rangle
\cdots |\gamma_N\rangle$.
When measured in the computational basis, the state given in \eref{eq: state}
collapses to $|\vec{\gamma}\rangle$ with a 
probability ${|b_{\vec{\gamma}}|}^2 = k_{\vec\gamma}$. 
The circuit in \fref{fig: circuit-pauli} uses this fact
to apply $\sigma_{\vec{\gamma}}$ on the main qubits
with a probability $k_{\vec\gamma}$ by using controlled operations 
conditioned on the state of the system being $|\vec{\gamma}\rangle$,
just as the Pauli channel is supposed to do.  
% 
% that apply $\sigma_{\vec{\gamma}}$ to the main qubits
% only if the ancilla qubits are in the state $|\vec{\gamma}\rangle$.
% 
% Therefore, the circuit from \fref{fig: circuit-pauli} 
% applies each operation $\sigma_{\vec{\gamma}}$ to the principal qubits 
% with a probability $k_{\vec{\gamma}}$, 
% just as a Pauli channel is supposed to do.

% \cpnote{Evaluar si al final ponemos algo al respecto de este comentario}
% \cpnote{Del circuito, parece ue las fases de las bs gammas no omportan mucho (o nada). 
% Esa libertad la hemos considerado? De alguna manera colapsan curvas de las que
% hemos visto?} \tbnote{Sí, es una libertad que se tiene y he considerado y significa que hay muchas elecciones de $b_{\gamma}$ para un mismo canal.
% Es una de las varias cosas que complicaba determinar uno a uno qué canales se pueden hacer.}
%

@@

For the particular case of a Pauli channel on one qubit, the circuit that
simulates it can be constructed as in \fref{fig: circuit-pauli-1},
which is a special case of \fref{fig: circuit-pauli}
but with all details  explicitly shown.
In said figure, the ancilla state
of \eref{eq: state} can be taken to be
$\sqrt{k_{00}} |00\rangle + \sqrt{k_{01}} |01\rangle + \sqrt{k_{10}} 
|10\rangle + \sqrt{k_{11}}|11\rangle$
and it is created on the ancilla qubits with the help of three rotations of angles
defined by the following equations:
\begin{align}
\label{eq:angle}
\cos\left(\dfrac{\theta_0}{2} \right) &= \sqrt{k_{00} + k_{01}},\\
\tan\left( \dfrac{\theta_1 + \theta_2}{2} \right) &= \sqrt{k_{01}/k_{00}}, \\
\tan\left( \dfrac{\theta_2 - \theta_1}{2} \right) &= \sqrt{k_{11}/k_{10}}.
\end{align}

\begin{figure} % {{{
\centering
\includegraphics{images/circuito_unqubit.pdf}
\caption{
{\bf One qubit Pauli channel circuit.} Circuit for a one-qubit Pauli channel,
which is a particular case of \fref{fig: circuit-pauli}.  Here we have two
ancilla qubits and we use three rotations
of angles given by \eref{eq:angle} to create the Ancilla State of
of \eref{eq: state} for the two ancilla qubits.
}
\label{fig: circuit-pauli-1} 
\end{figure} % }}}

@@

We took a sample of one-qubit Pauli channels
and evaluated their implementation on IBM's
ibmq-lima quantum computer~\cite{Qiskit},
as shown in \fref{fig:fids}. 
For each of the channels sampled, we used 
quantum process tomography~\cite{Qiskit, Chuang:1996} to obtain the 
operator $\xi_I$ corresponding to the implementation of the
circuit in the quantum computer. 
Then, we compared $\xi_I$ with the theoretical 
operator $\xi_T$ of the Pauli channel we wanted to implement.
To see how close the operators $\xi_I$ and 
$\xi_T$ are, we shall use the
diamond distance~\cite{Aharonov,Wilde}, which is defined by
\begin{equation}
||\xi_I - \xi_T ||_{\diamond}  = \max_{\rho} || (\xi_I \otimes I) \rho - (\xi_T \otimes I) \rho ||_1,
\end{equation}
with $I$ the identity map,
$|| \cdot ||_1$ the trace norm and the maximization
done over all density matrices $\rho$.
The calculation of this norm is done using the semi-definite program from reference \cite{Watrous}.
When the two channels are the same,
the diamond distance has a value of $0$, while in the case that the channels
are completely distinguishable, the distance reaches its maximum value of $2$~\cite{Benenti}.
For the analysis done in \fref{fig:fids}, we define a sort of ``diamond fidelity'' as: 
\begin{equation}
f = 1 - \dfrac{1}{2} ||\varepsilon_I - \varepsilon_T ||_{\diamond},
\label{eq:diamond-fid}
\end{equation}
which ranges from $0$, when the channels have a maximum distance, to  $1$,
when they are exactly equal.

\begin{figure} % {{{
\centering
\begin{subfigure}{0.9\textwidth}
\begin{subfigure}{.666\textwidth}
\centering
\includegraphics[width=.95\columnwidth]{images/tetra-def.png}
\end{subfigure}%
\begin{subfigure}{.333\textwidth}
\includegraphics[width=.95\columnwidth]{images/corte1-2.png} \\
\includegraphics[width=.95\columnwidth]{images/corte2-2.png} \\
\end{subfigure}
\includegraphics[width=.322\columnwidth]{images/corte3-2.png}
\includegraphics[width=.322\columnwidth]{images/corte4-2.png}
\includegraphics[width=.322\columnwidth]{images/corte5-2.png}
\end{subfigure}%
\begin{subfigure}{0.1\textwidth}
\includegraphics[width=.8\columnwidth]{images/col-2.png} \\ [2ex]
\end{subfigure}
\caption{Results of the diamond fidelities as defined in \eref{eq:diamond-fid}
for a sample of Pauli channels in the tetrahedron.  Notice that channels close
to the center of the tetrahedron have high fidelities, while those close to the
borders do not.  Moreover, we show the results for cuts of the tetrahedron at
different values of $\tau_3$.}
\label{fig:fids}
\end{figure} % }}}

Finally, using the representation of Pauli channels in a tetrahedron as in
\eref{eq: conditions-tetrahedron}, we show in \fref{fig:fids} the diamond
fidelity defined by \eref{eq:diamond-fid} for the channels analyzed.  We can
see that channels close to the completely depolarizing channel (that is close
to the center of the tetrahedron) have a high $f$, while those close to unitary
channels have much lower $f$.  This is reasonable because quantum computers are
prone to errors that depolarize qubits, which isn't very problematic when
trying to simulate depolarization but it is when simulating unitary processes.
Moreover, the algorithm of \fref{fig: circuit-pauli-1} is not optimal for
unitary channels (that is, the channels corresponding to the vertices of the
tetrahedron).  These straightforward channels could be accomplished more
efficiently by simply applying the corresponding Pauli operation directly.
Nevertheless, due to its general design to accommodate any Pauli channel, the
algorithm employs numerous quantum gates even in such scenarios.

@@

Just as Pauli channels, Pauli dynamical maps can be implemented using the circuit
of \fref{fig: circuit-pauli}. However, there is one difference: 
the state to be created on the ancilla qubits now 
depends on a parameter $p$, and it is represented by the expression:
\begin{eqnarray}
\label{eq: parametrized state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}}(p) |\vec{\gamma}\rangle.
\end{eqnarray}
Thus, we temporarily shift our focus from Pauli
channels and dynamical maps to the general problem of 
creating a circuit to generate a curve of
states like the one described in \eref{eq: parametrized state}.

In general, producing this curve of states for $N$ qubits will require
many rotations parametrized by $p$,
such as the three rotations used for the ancilla
qubits in \fref{fig: circuit-pauli-1}.
However, it would be preferable to achieve the same effect using only one parametrized rotation. 
This would allow us to interpret said rotation 
as a knob that smoothly traverses the curve of states.
Consequently, we are faced with the question of which curves of states, 
such as the one described in \eref{eq: parametrized state}, 
can be produced using just a single parametrized rotation. 
To clarify this, we provide the following definition 
for a circuit with one parametrized rotation.

@@

Based on this definition, we aim to determine which curves of 
states can be generated using 1PR circuits. 
To accomplish this, we begin by proving that all 1PR circuits have 
the form depicted in \fref{fig:1PR-circuit},
where the parametrized rotation is around $\sigma_3$ and
is applied to the last qubit.

\begin{figure} % {{{
\centering
\includegraphics{images/OPR-circuit.pdf}
\caption{\textbf{General form of a 1PR circuit}.
Any 1PR circuit can be transformed into this form,
where the rotation on the last qubit
can be controlled or not by any of the other qubits.
$A$ and $B$
are $N$-qubit gates that do not depend on the parameter $p$ and
$s= s(p)$ is a function of the parameter.
}
\label{fig:1PR-circuit}
\end{figure} % }}}

\begin{theorem}
An $N-$qubit 1PR circuit can always be transformed 
into the form shown in \fref{fig:1PR-circuit}.
\end{theorem}
\textbf{Proof:} 
First, we observe that according to the
definition, a 1PR circuit always consists of an operation $B$ followed 
by the parametrized rotation and then another operation $A$, 
where $A$ and $B$ are not parametrized.

@@

Therefore, a 1PR circuit can be transformed such
that the rotation is around $\sigma_3$ and is applied to the last qubit 
(possibly controlled by other qubits), 
resulting in the form depicted in \fref{fig:1PR-circuit}. 
$\blacksquare$ \\
$\;$\\

@@

This theorem implies that when starting from the state $|0\rangle$ 
or any other initial state, the only possible curves of states that can be 
created using a 1PR circuit are of the following form:
\begin{equation}
|\eta(p)\rangle = |c\rangle + e^{is(p)}|a\rangle + e^{-is(p)} |b\rangle,
\label{eq:curve-states}
\end{equation}
with conditions defined by the equations: 
\begin{equation}
\langle a| a\rangle + \langle b| b\rangle + \langle c| c\rangle = 1,  \quad
\langle a |b\rangle =
\langle a |c\rangle =
\langle b |c\rangle = 0.
\label{eq:conditions-vecs}
\end{equation}
% \cpnote{Es un poco confusa la notacion. Hablas de la norma de a con $|a|$, pero ahi parece
% que $a$ es un vector. Sugiero cambiar a $\langle a | a \rangle$ o algo asi sin cargar mucho
% la notacion. } \tbnote{Va, concuerdo, listo}
% \cpnote{Tambien siento qeu las palabras son un poco raras. Sugiero pasarlo por Chatgpt y 
% luego tunearlo (todo este parrafo). Avisame cuando quede eso!} \tbnote{Ya}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

@@

\linenumbers

@@

Such systems have been simulated because of their many applications, 
such as studying the emergence of multipartite entanglement~\cite{Andrea,Andrea_AD},
studying dissipative processes~\cite{Barreiro}
and modeling non-Markovian dynamics~\cite{Marsden}.
Among quantum systems, the simplest case is that of a qubit~\cite{chuangbook},
and withing them, the simplest class of channels that produce decoherence are
Pauli channels~\cite{geometry,Zbigniew,Davalos}. Indeed, they serve as
effective models for the noise affecting quantum devices~\cite{Flammia}.

To represent the algorithms implemented in quantum computers, either to
simulate a physical system or for some other purpose, one often uses a quantum
circuit~\cite{chuangbook}.  In this work, we shall also work with parametrized
quantum circuits, that is,  quantum circuits in which some of the operations
depend on variable parameters~\cite{cerezo}.  These circuits play an important
part in applications such as quantum machine learning~\cite{Benedetti} and
describing general quantum transformations dependent on parameters. Substantial
research has been devoted to enhancing the efficiency of these
circuits~\cite{Rasmussen}.

@@

% Intro  {{{

In this section  we introduce the concept of quantum channels, focusing on a
specific type called Pauli channels.  Furthermore, we define Pauli dynamical
maps, which are curves of Pauli channels parametrized by a variable.
% }}}
\subsection{Quantum channels} \label{subsec: Quantum Channels} % {{{

In quantum mechanics, a closed system's state is represented by a vector in a
Hilbert space $\mathcal{H}$.  The state's evolution is unitary and given by
Schrodinger's equation~\cite{Rieffel}.  However, in real-world situations,
quantum systems are usually open, which means that they interact with their
environment~\cite{breuer2007theory}.  For instance, the system's state may become
entangled with the environment, leading to a loss of information about the
system's state over time.

@@

% }}}
\subsection{Pauli channels}  \label{subsec: Pauli Channels} % {{{

@@

Pauli channels are some of the 
most fundamental noise models in quantum information science~\cite{Terhal}. 
Some notable examples of Pauli channels are the following:
\begin{itemize}
\item \textbf{Bit Flip Channel:} This is a channel that with probability $1-p$ leaves the qubit as it is 
and with probability $p$ applies the $\sigma_1$ matrix 
(which flips the basis states $|0\rangle$ and $|1\rangle$ of the qubit), and
so it is given by:
\begin{align*}
\mathcal{E}(\rho) = (1-p) \rho + p \sigma_1 \rho \sigma_1.
\end{align*}
Analogous channels exist using $\sigma_3$ (called the bit flip channel,
which has a probability $p$ of adding a relative phase $\pi$ to the state) or 
using $\sigma_2$ (called the phase flip channel, 
which has a probability $p$ of 
flipping the base states and also add a relative phase $\pi$).
\item \textbf{Depolarizing channel:} 
This channel has a probability $1-p$
of doing nothing to the qubit and a probability $p$ of converting it into the maximally mixed state $\dfrac{1}{2} \mathbb{I}$
and it can be written as:
\begin{equation}
\mathcal{E}(\rho) = (1-p)\rho + p \dfrac{1}{2} \mathbb{I} = \left(1 - \dfrac{3p}{4} \right) \sigma + \dfrac{p}{4} \sigma_1 \rho \sigma_1 + \dfrac{p}{4} \sigma_2 \rho \sigma_2+ \dfrac{p}{4} \sigma_3 \rho \sigma_3.
\end{equation} 
\end{itemize}

@@

These conditions imply that $(\tau_1,\tau_2,\tau_3)$
has to be inside a tetrahedron with
vertices $(1,1,1), (1,-1,-1), (-1,1,-1)$ and $(-1,-1,1)$. 
Therefore, the $\tau_{1,2,3}$ are numbers between 
$-1$ and $1$, which means that the components
$r_{\alpha}$ of the density matrix are always 
attenuated and possibly sign flipped.

@@

As in the one qubit case, Pauli channels for $N$ qubits 
attenuate the components $r_{\vec{\alpha}}$ of the density matrix.
This can be seen by substituting \eref{eq:  Density matrix Nqbit} in \eref{eq: pauli channel N-qubit}
and using the property of \eref{eq: propiedad-pauli}:
\begin{align*}
\mathcal{E}(\rho) &= \dfrac{1}{2^N} \sum_{\vec{\gamma},\vec{\alpha}} k_{\vec{\gamma}} r_{\vec{\alpha}} \sigma_{\vec{\gamma}} \sigma_{\vec{\alpha}} \sigma_{\vec{\gamma}} \\
& = \dfrac{1}{2^N} \sum_{\vec{\alpha}} \left( \sum_{\vec{\gamma}} {(A^{\otimes^N})}_{\vec{\alpha},\vec{\gamma}} \; k_{\vec{\gamma}} \right) r_{\vec{\alpha}} \sigma_{\vec{\alpha}},
\end{align*}
which means that applying the Pauli channel multiplies the 
components $r_{\vec{\alpha}}$ by $\tau_{\vec{\alpha}} := \sum_{\vec{\gamma}} {(A^{\otimes^N})}_{\vec{\alpha},\vec{\gamma}} k_{\vec{\gamma}}$.
% }}}
\subsection{Pauli dynamical maps} % {{{
\label{subsec: Pauli Dynamical Maps}

@@

In this section we propose a quantum circuit that simulates $N$-qubit Pauli
channels.  Moreover, we implement the circuit for $N=1$ on a quantum computer
and analyze the results using the diamond norm.
We find that close to the depolarizing channel, the general circuit simulator 
can implement such channels with the highest fidelity. 
% produce the highest 
% }}} 
\subsection{Description of the circuit for a Pauli channel} %{{{
\label{subsec: Description of the circuit}

To design the circuit that implements \eref{eq: pauli channel N-qubit}, we construct a state that 
includes the probabilities $k_{\gamma}$ on the ancilla qubits and subsequently
apply controlled Pauli operations on the main qubits.  The circuit that does
this is presented in \fref{Fig1}.

\begin{figure}[!h] % {{{
\centering
%\includegraphics{images/circuito_general.pdf}
\caption{{\bf Circuit for an $N$-qubit
Pauli channel.}
The circuit creates the state $\sum_{\vec{\gamma}} b_{\vec{\gamma}}|\vec{\gamma}\rangle$ 
on the $2N$ ancilla qubits denoted by $\tilde{q}$.}
\label{Fig1}
\end{figure} % }}}

The first part of the circuit involves the creation of the state
\begin{eqnarray}
\label{eq: state}
\text{Ancilla state =} 
\sum_{\vec{\gamma}} b_{\vec{\gamma}} |\vec{\gamma} \rangle
\end{eqnarray}
on the ancilla qubits, where $b_{\vec{\gamma}}$ are numbers such that $
{|b_{\vec{\gamma}}|}^2 = k_{\vec{\gamma}}$ and
the $2N$-qubit state $|\vec{\gamma}\rangle$ is defined as $|\gamma_1\rangle
\cdots |\gamma_N\rangle$.
When measured in the computational basis, the state given in \eref{eq: state}
collapses to $|\vec{\gamma}\rangle$ with a 
probability ${|b_{\vec{\gamma}}|}^2 = k_{\vec\gamma}$. 
The circuit in \fref{Fig1} uses this fact
to apply $\sigma_{\vec{\gamma}}$ on the main qubits
with a probability $k_{\vec\gamma}$ by using controlled operations 
conditioned on the state of the system being $|\vec{\gamma}\rangle$,
just as the Pauli channel is supposed to do.

@@

For the particular case of a Pauli channel on one qubit, the circuit that
simulates it can be constructed as in \fref{Fig2},
which is a special case of \fref{Fig1}
but with all details  explicitly shown.
In said figure, the ancilla state
of \eref{eq: state} can be taken to be
$\sqrt{k_{00}} |00\rangle + \sqrt{k_{01}} |01\rangle + \sqrt{k_{10}} 
|10\rangle + \sqrt{k_{11}}|11\rangle$
and it is created on the ancilla qubits with the help of three rotations of angles
defined by the following equations:
\begin{align}
\label{eq:angle}
\cos\left(\dfrac{\theta_0}{2} \right) &= \sqrt{k_{00} + k_{01}},\nonumber \\
\tan\left( \dfrac{\theta_1 + \theta_2}{2} \right) &= \sqrt{k_{01}/k_{00}},  \\
\tan\left( \dfrac{\theta_2 - \theta_1}{2} \right) &= \sqrt{k_{11}/k_{10}}.  \nonumber
\end{align}

\begin{figure}[!h] % {{{
\centering
%\includegraphics{images/circuito_unqubit.pdf}
\caption{
{\bf One-qubit Pauli channel circuit.} Circuit for a one-qubit Pauli channel,
which is a particular case of \fref{Fig1}.  Here we have two
ancilla qubits and we use three rotations
of angles given by \eref{eq:angle} to create the Ancilla State of
of \eref{eq: state} for the two ancilla qubits.
}
\label{Fig2} 
\end{figure} % }}}

@@

We took a sample of one-qubit Pauli channels
and evaluated their implementation on IBM's
ibmq-lima quantum computer~\cite{Qiskit},
as shown in \fref{Fig3}. 
For each of the channels sampled, we used 
quantum process tomography~\cite{Qiskit, Chuang:1996} to obtain the 
operator $\xi_I$ corresponding to the implementation of the
circuit in the quantum computer. 
Then, we compared $\xi_I$ with the theoretical 
operator $\xi_T$ of the Pauli channel we wanted to implement.
To see how close the operators $\xi_I$ and 
$\xi_T$ are, we shall use the
diamond distance~\cite{wildebook}, which is defined by
\begin{equation}
||\xi_I - \xi_T ||_{\diamond}  = \max_{\rho} || (\xi_I \otimes I) \rho - (\xi_T \otimes I) \rho ||_1,
\end{equation}
with $I$ the identity map,
$|| \cdot ||_1$ the trace norm and the maximization
done over all density matrices $\rho$.
The calculation of this norm is done using the semi-definite program from reference \cite{Watrous}.
When the two channels are the same,
the diamond distance has a value of $0$, while in the case that the channels
are completely distinguishable, the distance reaches its maximum value of $2$~\cite{Benenti}.
For the analysis done in \fref{Fig3}, we define a sort of ``diamond fidelity'' as: 
\begin{equation}
f = 1 - \dfrac{1}{2} ||\varepsilon_I - \varepsilon_T ||_{\diamond},
\label{eq:diamond-fid}
\end{equation}
which ranges from $0$, when the channels have a maximum distance, to  $1$,
when they are exactly equal.

\begin{figure}[!h] % {{{
\centering
\caption{{\bf Results of the diamond fidelities as defined in \eref{eq:diamond-fid}
for a sample of Pauli channels in the tetrahedron.}  Notice that channels close
to the center of the tetrahedron have high fidelities, while those close to the
borders do not.  Moreover, we show the results for cuts of the tetrahedron at
different values of $\tau_3$.}
\label{Fig3}
\end{figure} % }}}

Finally, using the representation of Pauli channels in a tetrahedron as in
\eref{eq: conditions-tetrahedron}, we show in \fref{Fig3} the diamond
fidelity defined by \eref{eq:diamond-fid} for the channels analyzed.  We can
see that channels close to the completely depolarizing channel (that is close
to the center of the tetrahedron) have a high $f$, while those close to unitary
channels have much lower $f$.  This is reasonable because quantum computers are
prone to errors that depolarize qubits, which isn't very problematic when
trying to simulate depolarization but it is when simulating unitary processes.
Moreover, the algorithm of \fref{Fig2} is not optimal for
unitary channels (that is, the channels corresponding to the vertices of the
tetrahedron).  These straightforward channels could be accomplished more
efficiently by simply applying the corresponding Pauli operation directly.
Nevertheless, due to its general design to accommodate any Pauli channel, the
algorithm employs numerous quantum gates even in such scenarios.

@@

Just as Pauli channels, Pauli dynamical maps can be implemented using the circuit
of \fref{Fig1}. However, there is one difference: 
the state to be created on the ancilla qubits now 
depends on a parameter $p$, and it is represented by the expression:
\begin{eqnarray}
\label{eq: parametrized state}
\sum_{\vec{\gamma}} b_{\vec{\gamma}}(p) |\vec{\gamma}\rangle.
\end{eqnarray}
Thus, we temporarily shift our focus from Pauli
channels and dynamical maps to the general problem of 
creating a circuit to generate a curve of
states like the one described in \eref{eq: parametrized state}.

In general, producing this curve of states for $N$ qubits will require
many rotations parametrized by $p$,
such as the three rotations used for the ancilla
qubits in \fref{Fig2}.
However, it would be preferable to achieve the same effect using only one parametrized rotation. 
This would allow us to interpret said rotation 
as a knob that smoothly traverses the curve of states.
Consequently, we are faced with the question of which curves of states, 
such as the one described in \eref{eq: parametrized state}, 
can be produced using just a single parametrized rotation. 
To clarify this, we provide the following definition 
for a circuit with one parametrized rotation.

@@

Based on this definition, we aim to determine which curves of 
states can be generated using 1PR circuits. 
To accomplish this, we begin by proving that all 1PR circuits have 
the form depicted in \fref{Fig4},
where the parametrized rotation is around $\sigma_3$ and
is applied to the last qubit.

\begin{figure}[!h] % {{{
\centering
%\includegraphics{images/OPR-circuit.pdf}
\caption{\textbf{General form of a 1PR circuit}.
Any 1PR circuit can be transformed into this form,
where the rotation on the last qubit
can be controlled or not by any of the other qubits.
$A$ and $B$
are $N$-qubit gates that do not depend on the parameter $p$ and
$s= s(p)$ is a function of the parameter.
}
\label{Fig4}
\end{figure} % }}}

\begin{theorem}
An $N-$qubit 1PR circuit can always be transformed 
into the form shown in \fref{Fig4}.
\end{theorem}
\textbf{Proof:} 
First, we observe that according to the
definition, a 1PR circuit always consists of an operation $B$ followed 
by the parametrized rotation and then another operation $A$, 
where $A$ and $B$ are not parametrized.

@@

Therefore, a 1PR circuit can be transformed such
that the rotation is around $\sigma_3$ and is applied to the last qubit 
(possibly controlled by other qubits), 
resulting in the form depicted in \fref{Fig4}. 
$\blacksquare$ \\
$\;$\\

@@

This theorem implies that when starting from the state $|0\rangle$ 
or any other initial state, the only possible curves of states that can be 
created using a 1PR circuit are of the following form:
\begin{equation}
|\eta(p)\rangle = |c\rangle + e^{is(p)}|a\rangle + e^{-is(p)} |b\rangle,
\label{eq:curve-states}
\end{equation}
with conditions defined by the equations: 
\begin{equation}
\langle a| a\rangle + \langle b| b\rangle + \langle c| c\rangle = 1,  \quad
\langle a |b\rangle =
\langle a |c\rangle =
\langle b |c\rangle = 0.
\label{eq:conditions-vecs}
\end{equation}

Moreover, it is possible to construct a 1PR circuit to generate any given curve of 
states described by \eref{eq:curve-states}. 
One approach to achieve this is by utilizing the circuit 
depicted in \fref{Fig4}, 
with the parametrized rotation $R$ applied to the last qubit controlled by all the 
other qubits. The operators $A$ and $B$ can be defined as follows:
\begin{align*}
&B|0\rangle = \sqrt{\langle a | a \rangle} |2^{N}-1 \rangle + \sqrt{\langle b | b \rangle} |2^{N}-2\rangle + \sqrt{\langle c |c \rangle} |2^N-3\rangle, \\
&A|2^N-1 \rangle = \frac{1}{\sqrt{\langle a | a \rangle}} |a\rangle,\; A|2^{N}-2\rangle = \frac{1}{\sqrt{\langle b | b \rangle}}|b\rangle, \; A|2^{N}-3\rangle = \frac{1}{\sqrt{\langle c | c \rangle}}|c\rangle.
\end{align*}
The remaining part of the operators $A$ and $B$ can be defined
in any arbitrary manner as long as they are unitary.
By starting from the initial state $|0\rangle$ and 
applying the circuit shown in \fref{Fig4}, straightforward calculations 
lead us to obtain the resulting curve of states described in \eref{eq:curve-states}.
